---
title: "Tidy Tuesday - Week 40"
date: "29/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyr)
library(dplyr)
library(lubridate)
library(readr)
library(ggplot2)
library(DT)
```

## Taylor Swift and Beyoncé

This document is to explore data visualisation of Week 40 of the Tidy Tuesday DataViz challenges.

```{r data}

#beyonce_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/beyonce_lyrics.csv')
#taylor_swift_lyrics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/taylor_swift_lyrics.csv')
sales <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/sales.csv')
charts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-29/charts.csv')



```

## Sales

Let's first take a look at the sales of their albums. There aren't that many so
the data will be a bit sparse. The TS data specifies world sales as 'WW' so 
renaming as 'World' to be comparable.

```{r sales}

# reformat dates
sales$released = as.Date(sales$released, format('%B %d, %Y'))
# and add month and year as separate cols
sales = sales %>% mutate(rel_month = format(released, "%b"))
sales = sales %>% mutate(rel_year = format(released, '%Y'))

# remove NA
sales = sales %>% filter(!is.na(sales))

# replace 'WW' with 'World'
sales[sales$country == 'WW', 'country'] <- 'World'

DT::datatable(sales)
```

It's a bit odd, but the 'Taylor Swift' album has no 'World' sales...

```{r plot1}

sales$rel_month = factor(sales$rel_month, levels = c( 'Dec', 'Nov', 'Oct', 'Sep', 'Aug', 'Jul', 'Jun', 'May', 'Apr', 'Mar', 'Feb','Jan'), ordered = TRUE)

sales_by_month = sales %>% 
  filter(country %in% c('UK', 'US', 'World')) %>%
  group_by(artist, rel_month, country) %>% 
  summarise(month_sales = sum(sales)/1000000)

sales_by_month = sales_by_month %>% complete(artist, country, rel_month)

ggplot(sales_by_month, aes(x = rel_month, y = month_sales, fill = artist)) +
  labs(x = 'Month of release',
       y = 'Total Sales (millions)') +
  geom_col(position = position_dodge(0.9), colour = 'darkgrey') +
  geom_text(aes(label = sprintf("%.1fm",round(month_sales, 1)), y = month_sales + 3.9), position = position_dodge(0.9), size = 3) +
  scale_x_discrete(limits = levels(sales$rel_month)) +
  scale_fill_manual(values = c('#fff44f','#af1946')) +
  ylim(c(0,30)) +
  coord_flip() + 
  facet_wrap(. ~ country) +
  theme_classic()
  
         

```

Sales outside of UK and US are very low and can't be seen on the full scale
so have been removed.

Taylor is 'Red' and Beyoncé is 'Lemon(ade)' for obvious reasons ;)

